This dissertation would not have been possible without the incredible patience of the nearly one hundred volunteers who participated in my experiments. I am grateful for their willingness to see a ten-hour experiment as a fun challenge and for the resolve of many of them to return over-and-over throughout my time at Stanford.

I've developed a wonderful habit during graduate school of glancing down the hall to see if Justin is in, or out. Not out of fear, fortunately, but because every day I'm excited to go and tell Justin about my successes and failures and my ideas. I have been extraordinarily lucky to have an advisor who still leaves their door open after five years of constant interruptions.

I have a deep gratitude for all of the vision science community at Stanford, in particular the faculty and students who have been a part of the Vision seminars and have helped guide my research in the last few years. Kalanit and Tony in particular watched out for me in my first year in a brand new lab and have continued to provide guidance at every stage of my research. Tirin's encouragement to connect my findings outside of the fMRI world has been critical to the final thesis and to my plans for the future. Finally, Bill Newsome's support (and wisdom) has been invaluable at every stage of data collection.

My views on vision science changed dramatically as part of a summer course in my fourth year and I am deeply thankful to the instructors and students who I shared my time with at Cold Spring Harbor. I hope to thank you each in person when we overlap. I want to extent a special thanks to Marlene Cohen, E.J. Chichilnisky, and Anne Urai for convincing me that you can study attention in other organisms than humans, I'm excited about the future path you have pushed me toward.

I have been lucky to be able to explore the mountains of California with other incredible scientists while doing my PhD. Andrew Lampinen, Mona Rosenke, and all of the other Stanford Psychology climbers have been a wonderful source of deep research conversations as well as incredible climbing partners.

I could not be here today without the support of my family and my climbing and life partner Allison. My parents and grandparents have made it easy for me to follow my interests wherever they lead.